Bow shocks, bow waves, and dust waves

### 2019-01-19 papers/Papers1234/sec-dust-bow-intro.tex
Finish appendix on CLoudy models

The internal pressure is the sum of wind ram pressure and the
effective radiation pressure that acts on the bow shell.  The
radiative momentum loss rate of the star is \(L/c\) and the wind
momentum loss rate can be expressed as
\begin{equation}
  \label{eq:wind-efficiency}
  \dot{M} V = \eta L / c \ , 
\end{equation}
where \(\eta\) is the momentum efficiency of the wind, which is typically
\(< 1\) \citep{Lamers:1999b}. If the optical depth is very large, then
all of the stellar radiative momentum, emitted with rate \(L/c\), is
trapped by the bow shell.  In the single scattering limit,\footnote{%
  Although it may seem inconsistent to assume single scattering in the
  case of high optical depths, this is defensible for the following
  reasons. (1)~The grain albedo is not that high (typically
  \(\sim 0.5\) at ultraviolet through optical wavelengths). (2)~The
  scattered radiation field is more isotropic than the stellar field,
  leading to cancellation in the radiative
  flux. (3)~Absorbed radiation is re-emitted at infrared
  wavelengths, where the dust opacity is very much lower.} %
and temporarily neglecting the wind, then pressure balance at the bow
apex, a distance \(R_0\) along the symmetry axis from the star is
given by
\begin{equation}
  \label{eq:rad-press-balance-thick}
  \frac{L}{4 \pi c R_0^2} = \rho v^2 \ ,
\end{equation}
which yields a fiducial bow shock radius in this optically thick limit
as
\begin{equation}
  \label{eq:Rstar}
  R_* = \left(\frac{L}{4\pi c \rho v^2}\right)^{1/2} \ .
\end{equation}

We now consider the opposite, optically thin limit.  If the total
opacity (gas plus dust) per total mass (gas plus dust) is \(\kappa\) (with
units of \si{cm^2.g^{-1}}), then the radiative acceleration is
\begin{equation}
  \label{eq:rad-accel}
  a_{\text{rad}} = \frac{\kappa L}{4 \pi c R^2} \ .
\end{equation}
Therefore, an incoming stream with initial velocity, \(v_\infty\), can be
brought to rest by radiation alone\footnote{%
  For simplicity, we here ignore the effects of gravity, which are
  important for low ratios of \(\kappa L / M\), see
  \S~\ref{sec:effects-gravity}.  We also ignore pressure gradients and
  shocks, which are important as the velocity approaches the sound
  speed, \(\sound\) (in \S~XX below, we show that the resultant
  corrections to \(R_0\) are of order \(\sound / v_\infty\)).} %
at a distance \(R\starstar\) where
\begin{equation}
  \label{eq:rad-poten}
  \int_{R\starstar}^\infty a_{\text{rad}} \, dr = \tfrac12 v_\infty^2 \ , 
\end{equation}
yielding
\begin{equation}
  \label{eq:rad:R0}
  R\starstar = \frac{\kappa L}{2\pi c v_\infty^2} \ .
\end{equation}
On the other hand, we can also argue as in the optically thick case
above by approximating the bow shell as a surface, and balancing
stellar radiation pressure against the ram pressure of the incoming
stream.  The important difference when the shell is not optically
thick is that only a fraction \(1 - e^{-\tau}\) of the radiative momentum
is absorbed by the bow, so that
equation~\eqref{eq:rad-press-balance-thick} is replaced with
\begin{equation}
  \label{eq:rad-press-balance-tau}
  \frac{L (1 - e^{-\tau})}{4 \pi c R_0^2} = \rho v^2 \ .
\end{equation}
In the optically thin limit, \(1 - e^{-\tau} \approx \tau\), so these two
descriptions can be seen to agree (\(R_0 \to R\starstar\)) so long as
\begin{equation}
  \label{eq:tau-thin}
  \tau = 2 \kappa \rho R_0 \ ,
\end{equation}
which we will assume to hold generally.

Then, defining a fiducial optical depth,
\begin{equation}
  \label{eq:tau-star}
  \tau_* = \rho \kappa R_* \ ,
\end{equation}
and adding in the stellar wind ram pressure\footnote{%
  We implicitly assume that the interaction of the stellar wind with
  the external stream can always be treated in the continuum limit.
  This will be true if either the collisional mean free path or the
  ion Larmor radius is much smaller than \(R_0\), which is almost
  always the case.} %
from equation~\eqref{eq:wind-efficiency}, we find that the general bow
radius can be written in terms of the fiducial radius as
\begin{equation}
  \label{eq:R0-definition}
  R_0 = x R_* \ ,
\end{equation}
where \(x\) is the solution of
\begin{equation}
  \label{eq:rad-full-x}
  x^2 - \bigl(1 - e^{-2 \tau_* x} \bigr) - \eta = 0 \ .
\end{equation}
Since this is a transcendental equation, \(x\) must be found
numerically, but we can write explicit expressions for three limiting
cases:
\begin{equation}
  \label{eq:x-cases}
  x \approx
  \begin{cases}
    \text{if \(\tau_* \gg 1\):} & (1 + \eta)^{1/2}  \\
    \text{if \(\tau_*^2 \ll 1\):} & \tau_* + \bigl( \tau_*^2 + \eta \bigr)^{1/2} \approx
    \begin{cases}
      \text{if \(\tau_*^2 \gg \eta\):} & 2 \tau_*  \\
      \text{if \(\tau_*^2 \ll \eta\):} & \eta^{1/2} 
    \end{cases}
  \end{cases}
\end{equation}
The first case, \(x \approx (1 + \eta)^{1/2}\), corresponds to a
\textit{radiation bow shock}; the second case,
\(x \approx 2 \tau_* \), corresponds to a \textit{radiation bow wave}; and the
third case, \(x \approx \eta^{1/2}\), corresponds to a \textit{wind bow shock}.
The two bow shock cases are similar in that the external stream is
oblivious to the presence of the star until it suddenly hits the bow
shock shell, differing only in whether it is radiation or wind that is
providing the internal pressure.  In the intermediate bow wave case,
on the other hand, the external stream is gradually decelerated by
absorption of photons as it approaches the bow.\footnote{A shock does
  still form in this case, but shocked material constitutes only a
  fraction of the total column density of the shell, see
  \S~\ref{sec:shape-bow-wave}.}

@@

stars, expressing stellar and ambient parameters in terms of typical
values as follows:

@@

which is nearly always small.  For clarity of exposition, we therefore
ignore \(\mathcal{A}\) in the following discussion, although it is
included in quantitative calculations.  The column density of the
shocked shell can be found, for example, from equations~(10) and~(12)
of \citet{Wilkin:1996a} in the limit \(v_\infty/V \to 0\) (Wilkin's parameter
\(\alpha\)) and \(\theta \to 0\).  This yields

@@

  \frac{n}{v_{10}^2} > \num{2.65e8} \, \frac{S_{49}^2 T_4^{3.4}}{L_4^3 (1 + \eta)^3} \ , 

@@

\citet{Sternberg:2003a}.  In the radiation bow wave case,
\(x = 2\tau_*\), and the condition can be written:

@@

@@

In Figure~\ref{fig:zones-v-n-plane} we show curves calculated from
equations~\eqref{eq:shock-n-jump} to~\eqref{eq:heating-coefficient},
corresponding to \(d\cool = R_0\) (thick blue line) and
\(d\cool = h_0\) (thin blue line), where \(h_0\) is the shell
thickness in the efficient cooling case.  In this context, \(n_0 = n\)
and \(n_2 = n\shell\), so that \(h_0\) follows from
equations~\eqref{eq:shocked-shell-column}
and~\eqref{eq:isothermal-shell-density} as
\begin{equation}
  \label{eq:strong-cooling-h0}
  h_0 = \tfrac34 \M_0^{-2} R_0 \ .
\end{equation}
The bends in the curves at \(v \approx \SI{50}{km.s^{-1}}\) are due to the
maximum in the cooling coefficient \(\Lambda(T)\) around
\(\SI{e5}{K}\).  For bows with outer stream densities above the thin
blue line, radiative cooling is so efficient that the bow shock can be
considered isothermal, and so the shell is dense and thin (at least,
in the apex region).  It can be seen that the ionization front
trapping always occurs at densities larger than this, which justifies
the use of equation~\eqref{eq:isothermal-shell-density} in the
previous section.  For bows with outer stream densities below the
thick blue line, cooling is unimportant and the bow shock can be
considered non-radiative.  In this case the shell is thicker than in
the radiative case,
\(h\shell/R_0 \approx \text{\numrange{0.2}{0.3}}\).\footnote{%
  An approximate value can be found from
  equation~\eqref{eq:shocked-shell-column} by substituting \(n = n_0\)
  and \(n\shell \approx n_1\), then using equation~\eqref{eq:shock-n-jump}.
  Consideration of the slight increase in density between the shock
  and the contact discontinuity reduces this value by 5--10\%.} %
For bows with outer stream densities between the two blue lines,
cooling does occur, albeit inefficiently, so that the shell thickness
is set by \(d\cool\) rather than \(h_0\).

@@

What happens to the dust grain following this catastrophic breakdown
of gas--grain coupling depends on the relation between the rip point
radius, \(R_\dag\), and the drag-free radiative turnaround radius,
\(R\starstar\).  If \(R_\dag > R\starstar\), then the grain's inertia
will still carry it in as far as \(R\starstar\) and the initial
trajectory will be almost identical to that described in
Appendix~\ref{sec:gas-free-bow} for the drag-free case. However, after
being turned around by the radiation field and pushed out past
\(R_\dag\) again, the grain will \emph{recouple} to the gas and be
dragged back for a second approach.\footnote{If the initial impact
  parameter of the trajectory is not strictly \(b = 0\), then the
  resultant lateral component of \(f\rad\) will mean that \(b\) will
  be much increased for the second approach.} %
We will refer to this as an \textit{inertia-confined dust wave}.  From
equations~(\ref{eq:Rdag-over-Rstar}, \ref{eq:Rstarstar-over-Rstar},
\ref{eq:dust-wave-high-density-condition}), the condition
\(R_\dag > R\starstar\) corresponds to
\(\tau_* < (\kappa/\kappa\grain) \tau_{*,\text{max}}\), which is indicated by dashed
lines in the left panel of Figure~\ref{fig:existence-dust-wave}.  If,
on the other hand, \(R_\dag < R\starstar\), then the tail wind provided
by the gas carries the grain closer to the star than its inertia would
naturally take it.  When the grain finally decouples at \(R_\dag\) it
experiences a much higher unbalanced \(f\rad\), which can initially
accelerate it to outward velocities significantly higher than the
inflow velocity if \(R_\dag \ll R\starstar\).  We will refer to this case
as a \textit{drag-confined dust wave}.

@@

\section{Magnetic effects on grain trajectories }
\label{sec:magn-effects-grain}

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/frozen-trajectories-multi}
  \caption{Sample grain trajectories for drag-confined dust waves with
    tight magnetic coupling.}
  \label{fig:frozen-trajectories}
\end{figure}

@@

The relative importance of the magnetic field can be characterized by
the ratio of the Larmor radius to the minimum radius, \(\Rmin\),
reached by the grain in the dust wave (see
\S~\ref{sec:grain-traj-along}), where \(\Rmin \approx R_\dag\) for
drag-confined dust waves, or \(\Rmin \approx R\starstar\) for
inertia-confined dust waves.  We write the field strength in terms of
the Alfvén speed,
\begin{equation}
  \label{eq:alfven}
  v\alfven = \frac{B}{(4\pi\rho\gas)^{1/2}}
  = 1.9\, \frac{B}{\si{\micro G}} n^{-1/2} \, \si{km.s^{-1}} \ ,
\end{equation}
and the grain charge \(z\grain e\) in terms of the potential \(\phi\) (eq.~[\ref{eq:phi-potential}]) to obtain
\begin{equation}
  \label{eq:larmor-over-Rdag}
  \frac{r\B}{R_\dag} = 0.0140 \,
  a_{\si{\um}}^2 \,
  \frac{w_\perp}{v\alfven} \,
  \left( \frac{\Xi_\dag}{L_4 T_4} \right)^{1/2}
  \frac{\rho\grain}{\phi_\dag}
\end{equation}
and
\begin{equation}
  \label{eq:larmor-over-Rstarstar}
  \frac{r\B}{R\starstar} = 0.0544 \,
  a_{\si{\um}}^3 \,
  \frac{w_\perp}{v\alfven} \,
  \frac{v_{10}^2 }{n^{1/2}} \,
  \frac{1}{L_4 T_4} \,
  \frac{\rho\grain^2}{\Qp \phi\starstar} \ ,
\end{equation}
where \(a_{\si{\um}} = a / \SI{1}{\um}\), \(\rho\grain\) is the grain
material density in \si{g.cm^{-3}}, and we have made use of
equations~(\ref{eq:Rstar-typical}, \ref{eq:Rdag-over-Rstar},
\ref{eq:dust-r0}).

If \(r\B / \Rmin \ll 1\), then the grains are so strongly coupled to the
field that they can be treated in the guiding-center approximation, in
which the trajectory is decomposed into a tight circular gyromotion
around the field lines, plus a sliding of the guiding center along the
field lines, which is governed by the radiation and drag forces (the
slow \(\bm{f} \times \bm{B}\) drift across the field lines is negligible in
this case, see App.~\ref{sec:tight-magn-coupl}).  In the opposite
limit, \(r\B / \Rmin \gg 1\), magnetic coupling is so weak that the
results of the previous sections are scarcely modified.  Assuming
\(w_\perp \sim v_\infty\) and adopting a threshold of
\(r\B / \Rmin < 0.1\), equations~(\ref{eq:larmor-over-Rdag},
\ref{eq:larmor-over-Rstarstar}) can be transformed into conditions on
the stream velocity (in \si{km.s^{-1}}) where tight magnetic coupling
will apply: \newcommand\freeze{\ensuremath{_{\text{tight}}}}
\begin{equation}
  \label{eq:velocity-strong-B-coupling}
  v_{\infty} < v\freeze \approx
  \begin{cases}
    \text{drag-confined:}
    & 0.8 \, a_{\si{\um}}^{-2} \, v\alfven \, L_4^{1/2}\\
    \text{inertia-confined:}
    & 6 \, a_{\si{\um}}^{-1} \, v\alfven^{1/3} \, n^{1/6} \, L_4^{1/3} \\
  \end{cases} \ ,
\end{equation}
where we have substituted typical values of the minor parameters
\(\Xi_\dag\), \(\phi_{\dag}\), \(\phi_{**}\), \(\rho\grain\),
\(T_4\).\footnote{%
  The most significant systematic variation in \(v\freeze\) from these
  suppressed parameters is due to grain composition, yielding slightly
  higher values for graphite than for silicate (\SI{+-0.15}{dex}).} %
It is apparent that \(v\freeze\) is very sensitive to the grain size.
For instance, taking a typical \hii{} region value of
\(v\alfven = \SI{2}{km.s^{-1}}\) \citep{Arthur:2011a} and
\(L_4 = 0.63\) (Tab.~\ref{tab:stars}, B1.5~V star), then for the
drag-confined case \(v\freeze \approx \SI{30}{km.s^{-1}}\) for \SI{0.2}{\um}
grains but \(v\freeze \approx \SI{3000}{km.s^{-1}}\) for \SI{0.02}{\um}
grains. Thus, for typical stream velocities of
\SIrange{20}{100}{km.s^{-1}}, the small grains are always tightly
coupled to the magnetic field, but the large grains are only loosely
coupled for the faster streams.

>>>>

The internal pressure is the sum of wind ram pressure and the
effective radiation pressure that acts on the bow shell.  The
radiative momentum loss rate of the star is \(L/c\) and the wind
momentum loss rate can be expressed as
\begin{equation}
  \label{eq:wind-efficiency}
  \dot{M} V = \eta L / c \ , 
\end{equation}
where \(\eta\) is the momentum efficiency of the wind, which is typically
\(< 1\) \citep{Lamers:1999b}. If the optical depth is very large, then
all of the stellar radiative momentum, emitted with rate \(L/c\), is
trapped by the bow shell.  In the single scattering limit,\footnote{%
  Although it may seem inconsistent to assume single scattering in the
  case of high optical depths, this is defensible for the following
  reasons. (1)~The grain albedo is not that high (typically
  \(\sim 0.5\) at ultraviolet through optical wavelengths). (2)~The
  scattered radiation field is more isotropic than the stellar field,
  leading to cancellation in the radiative flux. (3)~Absorbed
  radiation is re-emitted at infrared wavelengths, where the dust
  opacity is very much lower.} %
and temporarily neglecting the wind for clarity of exposition, then
pressure balance at the bow apex, a distance \(R_0\) along the
symmetry axis from the star is given by
\begin{equation}
  \label{eq:rad-press-balance-thick}
  \frac{L}{4 \pi c R_0^2} = \rho v^2 \ ,
\end{equation}
which yields a fiducial bow shock radius in this optically thick,
radiation-only limit as
\begin{equation}
  \label{eq:Rstar}
  R_* = \left(\frac{L}{4\pi c \rho v^2}\right)^{1/2} \ .
\end{equation}

We now consider the opposite, optically thin limit.  If the total
opacity (gas plus dust) per total mass (gas plus dust) is \(\kappa\) (with
units of \si{cm^2.g^{-1}}), then the radiative acceleration is
\begin{equation}
  \label{eq:rad-accel}
  a_{\text{rad}} = \frac{\kappa L}{4 \pi c R^2} \ .
\end{equation}
Therefore, an incoming stream with initial velocity, \(v_\infty\), can be
brought to rest by radiation alone
% \footnote{%
%   For simplicity, we here ignore the effects of gravity, which are
%   important for low ratios of \(\kappa L / M\), see
%   \S~\ref{sec:effects-gravity}.  We also ignore pressure gradients and
%   shocks, which are important as the velocity approaches the sound
%   speed, \(\sound\) (in \S~XX below, we show that the resultant
%   corrections to \(R_0\) are of order \(\sound / v_\infty\)).} %
at a distance \(R\starstar\) where
\begin{equation}
  \label{eq:rad-poten}
  \int_{R\starstar}^\infty a_{\text{rad}} \, dr = \tfrac12 v_\infty^2 \ , 
\end{equation}
yielding
\begin{equation}
  \label{eq:rad:R0}
  R\starstar = \frac{\kappa L}{2\pi c v_\infty^2} \ .
\end{equation}
On the other hand, we can also argue as in the optically thick case
above by approximating the bow shell as a surface, and balancing
stellar radiation pressure against the ram pressure of the incoming
stream.  The important difference when the shell is not optically
thick is that only a fraction \(1 - e^{-\tau}\) of the radiative momentum
is absorbed by the bow, so that
equation~\eqref{eq:rad-press-balance-thick} is replaced with
\begin{equation}
  \label{eq:rad-press-balance-tau}
  \frac{L (1 - e^{-\tau})}{4 \pi c R_0^2} = \rho v^2 \ .
\end{equation}
In the optically thin limit, \(1 - e^{-\tau} \approx \tau\), so these two
descriptions can be seen to agree (\(R_0 \to R\starstar\)) so long as
\begin{equation}
  \label{eq:tau-thin}
  \tau = 2 \kappa \rho R_0 \ ,
\end{equation}
which we will assume to hold generally.

Then, defining a fiducial optical depth,
\begin{equation}
  \label{eq:tau-star}
  \tau_* = \rho \kappa R_* \ ,
\end{equation}
and now reinstating the stellar wind ram pressure term
% \footnote{%
%   We implicitly assume that the interaction of the stellar wind with
%   the external stream can always be treated in the continuum limit.
%   This will be true if either the collisional mean free path or the
%   ion Larmor radius is much smaller than \(R_0\), which is almost
%   always the case.} %
from equation~\eqref{eq:wind-efficiency}, we find that the general bow
radius can be written in terms of the fiducial radius as
\begin{equation}
  \label{eq:R0-definition}
  R_0 = x R_* \ ,
\end{equation}
where \(x\) is the solution of
\begin{equation}
  \label{eq:rad-full-x}
  x^2 - \bigl(1 - e^{-2 \tau_* x} \bigr) - \eta = 0 \ .
\end{equation}
Since this is a transcendental equation, \(x\) must be found
numerically, but we can write explicit expressions for three limiting
cases:
\begin{equation}
  \label{eq:x-cases}
  x \approx
  \begin{cases}
    \text{if \(\tau_* \gg 1\):} & (1 + \eta)^{1/2}  \\
    \text{if \(\tau_*^2 \ll 1\):} & \tau_* + \bigl( \tau_*^2 + \eta \bigr)^{1/2} \approx
    \begin{cases}
      \text{if \(\tau_*^2 \gg \eta\):} & 2 \tau_*  \\
      \text{if \(\tau_*^2 \ll \eta\):} & \eta^{1/2} 
    \end{cases}
  \end{cases}
\end{equation}
The first case, \(x \approx (1 + \eta)^{1/2}\), corresponds to a
\textit{radiation bow shock} (RBS); the second case,
\(x \approx 2 \tau_* \), corresponds to a \textit{radiation bow wave} (RBW);
and the third case, \(x \approx \eta^{1/2}\), corresponds to a \textit{wind bow
  shock} (WBS).  The two bow shock cases are similar in that the
external stream is oblivious to the presence of the star until it
suddenly hits the bow shock shell, differing only in whether it is
radiation or wind that is providing the internal pressure.  In the
intermediate bow wave case, on the other hand, the external stream is
gradually decelerated by absorption of photons as it approaches the
bow.\footnote{A shock can still form in this case, but shocked
  material constitutes only a small fraction of the total column
  density of the shell.}

\subsection{Dependence on stellar type}
\label{sec:depend-stell-type}

@@

stars, as well as cool and hot supergiants, expressing stellar and
ambient parameters in terms of typical values as follows:

@@

\citet{Sternberg:2003a}.  Since \(\mathcal{A} \ll \mathcal{F}\) in
nearly all cases, for clarity of exposition we ignore \(\mathcal{A}\)
in the following discussion, although it is included in quantitative
calculations.  The column density of the shocked shell can be found,
for example, from equations~(10) and~(12) of \citet{Wilkin:1996a} in
the limit \(v_\infty/V \to 0\) (Wilkin's parameter \(\alpha\)) and
\(\theta \to 0\).  This yields

@@

  \text{RBS:}\quad
  \frac{n}{v_{10}^2} > \num{2.65e8} \, \frac{S_{49}^2 T_4^{3.4}}{L_4^3 (1 + \eta)^3} \ .

@@

In the radiation bow wave case, \(x = 2\tau_*\), and the condition can be
written:

@@

  \text{RBW:}\quad

@@

In Figure~\ref{fig:zones-v-n-plane} we show curves calculated from
equations~\eqref{eq:shock-n-jump} to~\eqref{eq:heating-coefficient},
corresponding to \(d\cool = R_0\) (thick blue line) and
\(d\cool = h_0\) (thin blue line), where \(h_0\) is the shell
thickness in the efficient cooling case.  In this context, \(n_0 = n\)
and \(n_2 = n\shell\), so that \(h_0\) follows from
equations~\eqref{eq:shocked-shell-column}
and~\eqref{eq:isothermal-shell-density} as
\begin{equation}
  \label{eq:strong-cooling-h0}
  h_0 = \tfrac34 \M_0^{-2} R_0 \ .
\end{equation}
The bends in the curves at \(v \approx \SI{50}{km.s^{-1}}\) are due to the
maximum in the cooling coefficient \(\Lambda(T)\) around
\(\SI{e5}{K}\).  For bows with outer stream densities above the thin
blue line, radiative cooling is so efficient that the bow shock can be
considered isothermal, and so the shell is dense and thin (at least,
in the apex region).  It can be seen that the ionization front
trapping always occurs at densities larger than this, which justifies
the use of equation~\eqref{eq:isothermal-shell-density} in the
previous section.  For bows with outer stream densities below the
thick blue line, cooling is unimportant and the bow shock can be
considered non-radiative.  In this case the shell is thicker than in
the radiative case,
\(h\shell/R_0 \approx \text{\numrange{0.2}{0.3}}\).\footnote{%
  An approximate value can be found from
  equation~\eqref{eq:shocked-shell-column} by substituting \(n = n_0\)
  and \(n\shell \approx n_1\), then using equation~\eqref{eq:shock-n-jump}.
  Consideration of the slight increase in density between the shock
  and the contact/tangential discontinuity reduces this value by 5--10\%.} %
For bows with outer stream densities between the two blue lines,
cooling does occur, albeit inefficiently, so that the shell thickness
is set by \(d\cool\) rather than \(h_0\).

@@

What happens to the dust grain following this catastrophic breakdown
of gas--grain coupling depends on the relation between the rip point
radius, \(R_\dag\), and the drag-free radiative turnaround radius,
\(R\starstar\).  If \(R_\dag > R\starstar\), then the grain's inertia
will still carry it in as far as \(R\starstar\) and the initial
trajectory will be almost identical to that described in
Appendix~\ref{sec:gas-free-bow} for the drag-free case. However, after
being turned around by the radiation field and pushed out past
\(R_\dag\) again, the grain will \emph{recouple} to the gas and be
dragged back for a second approach.\footnote{If the initial impact
  parameter of the trajectory is not strictly \(b = 0\), then the
  resultant lateral component of \(f\rad\) will mean that \(b\) will
  be much increased for the second approach.} %
We will refer to this as an \textit{inertia-confined dust wave}
(IDW).  From equations~(\ref{eq:Rdag-over-Rstar},
\ref{eq:Rstarstar-over-Rstar},
\ref{eq:dust-wave-high-density-condition}), the condition
\(R_\dag > R\starstar\) corresponds to
\(\tau_* < (\kappa/\kappa\grain) \tau_{*,\text{max}}\), which is indicated by dashed
lines in the left panel of Figure~\ref{fig:existence-dust-wave}.  If,
on the other hand, \(R_\dag < R\starstar\), then the tail wind provided
by the gas carries the grain closer to the star than its inertia would
naturally take it.  When the grain finally decouples at \(R_\dag\) it
experiences a much higher unbalanced \(f\rad\), which can initially
accelerate it to outward velocities significantly higher than the
inflow velocity if \(R_\dag \ll R\starstar\).  We will refer to this case
as a \textit{drag-confined dust wave} (DDW).

@@

\section{Magnetic coupling of grains}
\label{sec:magn-effects-grain}

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/frozen-trajectories-multi}
  \caption{Sample grain trajectories for drag-confined dust waves with
    tight magnetic coupling and a quasi-parallel field
    orientation. These are the same models as in the upper row of
    Fig.~\ref{fig:frozen-stream}. (a)~Incident stream velocity of
    \(v_\infty = \SI{40}{km.s^{-1}}\), showing quasi-limit-cycle behavior
    (\numrange{0}{4000} years).  (b)~Incident stream velocity of
    \(v_\infty = \SI{80}{km.s^{-1}}\), showing quasi-stagnation behavior
    (\numrange{500}{2000} years).  In both cases, the streamline with
    initial impact parameter \(y = -0.4 R_\dag\) is shown.}
  \label{fig:frozen-trajectories}
\end{figure}

@@

The relative importance of the magnetic field can be characterized by
the ratio of the Larmor radius to the minimum radius, \(\Rmin\),
reached by the grain in the dust wave (see
\S~\ref{sec:grain-traj-along}), where \(\Rmin \approx R_\dag\) for
drag-confined dust waves (DDW), or \(\Rmin \approx R\starstar\) for
inertia-confined dust waves (IDW).  We write the field strength in
terms of the Alfvén speed,
\begin{equation}
  \label{eq:alfven}
  v\alfven = \frac{B}{(4\pi\rho\gas)^{1/2}}
  = 1.9\, \frac{B}{\si{\micro G}} n^{-1/2} \, \si{km.s^{-1}} \ ,
\end{equation}
and the grain charge \(z\grain e\) in terms of the potential \(\phi\) (eq.~[\ref{eq:phi-potential}]) to obtain
\begin{equation}
  \label{eq:larmor-over-Rdag}
  \text{DDW:}\quad \frac{r\B}{\Rmin} =  
  \frac{r\B}{R_\dag} = 0.0140 \,
  a_{\si{\um}}^2 \,
  \frac{w_\perp}{v\alfven} \,
  \left( \frac{\Xi_\dag}{L_4 T_4} \right)^{1/2}
  \frac{\rho\grain}{\phi_\dag}
\end{equation}
and
\begin{equation}
  \label{eq:larmor-over-Rstarstar}
  \text{IDW:}\quad \frac{r\B}{\Rmin} =  
  \frac{r\B}{R\starstar} = 0.0544 \,
  a_{\si{\um}}^3 \,
  \frac{w_\perp}{v\alfven} \,
  \frac{v_{10}^2 }{n^{1/2}} \,
  \frac{1}{L_4 T_4} \,
  \frac{\rho\grain^2}{\Qp \phi\starstar} \ ,
\end{equation}
where \(a_{\si{\um}} = a / \SI{1}{\um}\), \(\rho\grain\) is the grain
material density in \si{g.cm^{-3}}, and we have made use of
equations~(\ref{eq:Rstar-typical}, \ref{eq:Rdag-over-Rstar},
\ref{eq:dust-r0}).

If \(r\B / \Rmin \ll 1\), then the grains are so strongly coupled to the
field that they can be treated in the guiding-center approximation, in
which the trajectory is decomposed into a tight circular gyromotion
around the field lines, plus a sliding of the guiding center along the
field lines, which is governed by the radiation and drag forces (the
slow \(\bm{f} \times \bm{B}\) drift across the field lines is negligible in
this case, see App.~\ref{sec:tight-magn-coupl}).  In the opposite
limit, \(r\B / \Rmin \gg 1\), magnetic coupling is so weak that the
non-magnetic results of \S~\ref{sec:imperf-coupl-betw} are scarcely
modified.  Assuming \(w_\perp \sim v_\infty\) and adopting a threshold of
\(r\B / \Rmin < 0.1\), equations~(\ref{eq:larmor-over-Rdag},
\ref{eq:larmor-over-Rstarstar}) can be transformed into conditions on
the stream velocity (in \si{km.s^{-1}}) where tight magnetic coupling
will apply: \newcommand\freeze{\ensuremath{_{\text{tight}}}}
\begin{equation}
  \label{eq:velocity-strong-B-coupling}
  v_{\infty} < v\freeze \approx
  \begin{cases}
    \text{drag-confined:}
    & 0.8 \, a_{\si{\um}}^{-2} \, v\alfven \, L_4^{1/2}\\
    \text{inertia-confined:}
    & 6 \, a_{\si{\um}}^{-1} \, v\alfven^{1/3} \, n^{1/6} \, L_4^{1/3} \\
  \end{cases} \ ,
\end{equation}
where we have substituted typical values of the minor parameters
\(\Xi_\dag\), \(\phi_{\dag}\), \(\phi_{**}\), \(\rho\grain\),
\(T_4\).\footnote{%
  The most significant systematic variation in \(v\freeze\) from these
  suppressed parameters is due to grain composition, yielding slightly
  higher values for graphite than for silicate (\SI{+-0.15}{dex}).} %
It is apparent that \(v\freeze\) is very sensitive to the grain size.
For instance, taking a typical \hii{} region value of
\(v\alfven = \SI{2}{km.s^{-1}}\) \citep{Arthur:2011a} and
\(L_4 = 0.63\) (Tab.~\ref{tab:stars}, B1.5~V star), then for the
drag-confined case \(v\freeze \approx \SI{30}{km.s^{-1}}\) for \SI{0.2}{\um}
grains but \(v\freeze \approx \SI{3000}{km.s^{-1}}\) for \SI{0.02}{\um}
grains. Thus, for typical stream velocities of
\SIrange{20}{100}{km.s^{-1}}, the small grains are always tightly
coupled to the magnetic field, but the large grains are only loosely
coupled for the faster streams.

\subsection{Grain trajectories with tight magnetic coupling}
\label{sec:grain-traj-with}
